Searching for Software in NSF Awards

### 2023-02-10 paper/paper.qmd
Introduction done

@@

## Introduction

Some content!

## Data Collection and Annotation {#sec-data-collection}

### Finding Software Produced by NSF Awards {#sec-finding-soft}

@@

### Software Production Annotation

@@

### Linking GitHub Repositories to NSF Awards

@@

## Predictive Models

@@

from_abstract
```

@@

from_po
```

@@

## The NSF-Soft-Search Dataset

@@

    desc_table_rows.append({
        "Funding Program": program,
        "# Awards Funded": total_awards_funded,
        "# Awards 'Produce Software'": num_awards_will_produce_software,
        "% Awards 'Produce Software'": percent_awards_will_produce,
    })

desc_table = pd.DataFrame(desc_table_rows).sort_values(
    by=["# Awards Funded", "% Awards 'Produce Software'"],
    ascending=False,
).reset_index(drop=True)
desc_table
```

### Trends and Observations

* Trends in software production over time from abstract
```{python}
#| code-fold: true
#| warning: false

@@

        soft_produced_over_time_rows.append({
            "Funding Program": program_name,
            "Award Start Year": start_year,
            "% Awards 'Produce Software'": predicted_count / award_count,
        })

@@

# Plot over time
sns.barplot(
    data=soft_produced_over_time,
    x="Award Start Year",
    y="% Awards 'Produce Software'",
    hue="Funding Program",
    order=[
        str(i)
        for i in range(
            high_soft_prog_df.startYear.astype(int).min(),
            high_soft_prog_df.startYear.astype(int).max(),
        )
    ],
)
plt.xticks(rotation=45, ha="right")
plt.tight_layout()
```

@@

        soft_produced_over_duration_rows.append({
            "Funding Program": program_name,
            "Award Duration (Years)": award_duration,
            "% Awards 'Produce Software'": predicted_count / award_count,
        })

@@

# Plot over time
sns.barplot(
    data=soft_produced_over_duration,
    x="Award Duration (Years)",
    y="% Awards 'Produce Software'",
    hue="Funding Program",
)
plt.tight_layout()
```

@@

## Conclusion

We introduce NSF-Soft-Search, a pair of novel datasets for studying software production from NSF funded projects. The NSF-Soft-Search GitHub Annotations dataset is used to train models which aim to predict software production from either the NSF award abstract text or the project outcomes report text. We used these models to generate the NSF-Soft-Search Inferred dataset. The NSF-Soft-Search Inferred dataset includes project metadata, abstract, project outcomes report, and a predictions of software production for each NSF funded project between 2010 and 2023. We hope that NSF-Soft-Search helps further new studies and findings in understanding the role software development plays in scholarly publication.

@@

### Limitations

@@

### Future Work

@@

## Acknowledgements

@@

::: {.content-visible unless-format="html"}
## References
:::

>>>>

## Number sections (required for section cross ref)
number-sections: true

@@

# Introduction

Software production, use, and reuse is an increasingly crucial part of scholarly work [@open_source_code_repo_predict_impact;@Trisovic2021ALS]. While historically ignored, citing and referencing software used during the course of research is becoming common with new standards for software citation [@Katz2021RecognizingTV;@Du2022UnderstandingPI] and work in extracting software references in existing literature [@cz_software_mentions]. However, records of software production are not available at scale the way other bibliographic metrics are. To solve this problem, we introduce two related datasets for studying and inferring software produced as a part of research, which we refer to as the NSF-Soft-Search dataset.

The NSF-Soft-Search dataset is aimed at identifying research projects which are likely to have produced software during the course of a project. We start by identifying GitHub repositories which acknowledge funding from at least one National Science Foundation (NSF) award. We then annotate each GitHub repository found with a binary decision for it's contents: software or not-software (research notes, course materials, etc.). We link each annotated GitHub repository to the specific NSF award ID(s) referenced in it's README.md file. Finally, we compile the NSF-Soft-Search GitHub Annotations dataset using the annotations for each GitHub repository, and the text from the linked NSF award abstract and the project outcomes report.

Using the NSF-Soft-Search GitHub Annotations dataset, we train a variety of models which aim to predict software production using either the NSF award abstract or project outcomes report text as input. We use the best performing models to then infer software production against all awards funded by the National Science Foundation from 2010 to 2023. The predictions and metadata for each NSF award between the 2010 and 2023 period are compiled to form the NSF-Soft-Search Inferred dataset.

The rest of the paper proceeds as follows. In @sec-data-collection we detail the data collection and annotation process used for creating the NSF-Soft-Search GitHub Annotations dataset. In @sec-models we briefly describe the model training process and report results for the all of the models trained. In @sec-nsf-soft-search-dataset we provide summary statistics for the NSF-Soft-Search Inferred dataset and use the dataset to observe trends in software production over time. We conclude with discussion regarding the limitations of our approach and opportunities for future work.

# Data Collection and Annotation {#sec-data-collection}

## Finding Software Produced by NSF Awards {#sec-finding-soft}

@@

## Software Production Annotation

@@

## Linking GitHub Repositories to NSF Awards

@@

# Predictive Models {#sec-models}

@@

Markdown(from_abstract.to_markdown(index=False))
```

@@

Markdown(from_po.to_markdown(index=False))
```

@@

# The NSF-Soft-Search Dataset {#sec-nsf-soft-search-dataset}

@@

    desc_table_rows.append({
        "Funding Program": program,
        "# Awards Funded": total_awards_funded,
        "# 'Produce Software'": num_awards_will_produce_software,
        "% 'Produce Software'": percent_awards_will_produce,
    })

desc_table = pd.DataFrame(desc_table_rows).sort_values(
    by=["# Awards Funded", "% 'Produce Software'"],
    ascending=False,
).reset_index(drop=True)
Markdown(desc_table.to_markdown(index=False))
```

## Trends and Observations

@@

        soft_produced_over_time_rows.append({
            "Funding Program": program_name,
            "Award Start Year": start_year,
            "% 'Produce Software'": predicted_count / award_count,
        })

@@

# Ensure that hue colors are consistent throughout plots
soft_produced_over_time = soft_produced_over_time.sort_values(
  by="Funding Program",
)

# Plot over time
ax = sns.barplot(
    data=soft_produced_over_time,
    x="Award Start Year",
    y="% 'Produce Software'",
    hue="Funding Program",
    order=[
        str(i)
        for i in range(
            high_soft_prog_df.startYear.astype(int).min(),
            high_soft_prog_df.startYear.astype(int).max(),
        )
    ],
)
sns.move_legend(
    ax,
    "lower center",
    bbox_to_anchor=(.5, 1),
    ncol=len(soft_produced_over_time["Funding Program"].unique()),
    title="Funding Program",
    frameon=False,
)
plt.xticks(rotation=45, ha="right")
plt.tight_layout()
```

@@

        soft_produced_over_duration_rows.append({
            "Funding Program": program_name,
            "Award Duration (Years)": award_duration,
            "% 'Produce Software'": predicted_count / award_count,
        })

@@

# Ensure that hue colors are consistent throughout plots
soft_produced_over_duration = soft_produced_over_duration.sort_values(
  by="Funding Program",
)

# Plot over time
ax = sns.barplot(
    data=soft_produced_over_duration,
    x="Award Duration (Years)",
    y="% 'Produce Software'",
    hue="Funding Program",
)
sns.move_legend(
    ax,
    "lower center",
    bbox_to_anchor=(.5, 1),
    ncol=len(soft_produced_over_time["Funding Program"].unique()),
    title="Funding Program",
    frameon=False,
)
plt.tight_layout()
```

@@

# Conclusion

We introduce NSF-Soft-Search, a pair of novel datasets for studying software production from NSF funded projects. The NSF-Soft-Search GitHub Annotations dataset is used to train models which aim to predict software production from either the NSF award abstract text or the project outcomes report text. We used these models to generate the NSF-Soft-Search Inferred dataset. The NSF-Soft-Search Inferred dataset includes project metadata, abstract, project outcomes report, and predictions of software production for each NSF funded project between 2010 and 2023. We hope that NSF-Soft-Search helps further new studies and findings in understanding the role software development plays in scholarly publication.

@@

## Limitations

@@

## Future Work

@@

# Acknowledgements

@@

::: {.content-visible unless-format="html"}
# References
:::
